Trade-offs among plant floral and reproductive traits determine interactions with floral visitors

### 2021-08-11 Draft/Manuscript.tex
manuscript

\documentclass[11pt,a4paper,]{article}
\usepackage[]{mathpazo}
\usepackage{setspace}
\setstretch{2}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{
            pdftitle={Plant reproductive strategies structure plant-pollinator networks},
            pdfauthor={Jose B. Lanuza1,2*; Romina Rader1; Jamie Stavert3; Liam K. Kendall4; Manu E. Saunders1; Ignasi Bartomeus2},
            colorlinks=true,
            linkcolor=RoyalBlue,
            citecolor=Blue,
            urlcolor=RoyalBlue,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

% Add Supplementary Tables and Figures
% Code from https://stackoverflow.com/a/51337664
\newcommand{\beginsupplement}{
  \setcounter{table}{0}  
  \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} 
  \renewcommand{\thefigure}{S\arabic{figure}}
}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\title{Plant reproductive strategies structure plant-pollinator networks}
\author{Jose B. Lanuza\textsuperscript{1,2}* \and Romina Rader\textsuperscript{1} \and Jamie Stavert\textsuperscript{3} \and Liam K. Kendall\textsuperscript{4} \and Manu E. Saunders\textsuperscript{1} \and Ignasi Bartomeus\textsuperscript{2}}
\date{}

@@

\vspace{1cm} \hrule

\section{Abstract}\label{abstract}

Increasing availability of large datasets have improved our
understanding of organismal life history and pollination biology.
However, most attention has focused on vegetative traits when delimiting
plant life strategies and overlooked key floral and reproductive traits.
In addition, plant-pollinator research rarely considers the reproductive
biology of the species and has overemphasized the study of highly
specialized interactions. Greater knowledge of the plant reproductive
trade-offs between these traits can shed light on the macroecological
patterns and help our understanding of life strategies.

Here, we compiled 20 reproductive traits for 1,506 plant species from 28
plant-pollinator network studies and 18 different countries that
constitute a total of 64 unique networks and 8 meta-webs. We explore the
different trade-offs of floral, reproductive, and vegetative traits of
the main quantitative traits in the trait space. Then, we evaluate
whether the frequency of pollinator visits within networks change across
the primary axes of trait variation and pollinator functional groups and
how individual traits contribute to the overall visitation patterns.
Finally, we investigate if the functional role of the plant species
(interaction frequency, normalized degree and specialization) in the
network is associated with the primary axes of trait variation and the
different plant traits.

We found that 51.8\% of the trait variation was explained by two
independent axes that determined plant form and function. The first axis
showed the flower number versus flower size compromise (PC1, 26.72\%)
where species were divided in two opposed strategies with high
investment in flower number and plant height on one extreme or high
investment in flower size, style length and larger number of ovules on
the other. The second axis showed the pollinator dependence trade-off
(PC2, 25.08\%) between species with high pollinator dependence with
greater plant height and higher investment in floral traits on one end
and low pollinator dependence with smaller height and lower investment
in floral traits on the other. An additional principal component
analysis with the subset of species with quantitative data of pollen and
nectar showed that nectar quantity was positively associated with flower
size, style length and ovule number (PC1, 24.70\%) and pollen quantity
was positively correlated with flower number and plant height and
negatively associated with autonomous selfing (PC2, 20.59\%). The main
axes of trait variation with the full set of species did not explain the
overall visitation patterns but showed marked trends for the different
functional groups. For instance, all pollinator functional groups had
higher frequency of visits on plant species with high pollinator
dependence (PC2); and bees (Anthophila), the most abundant functional
group, had higher frequency of visits on species with high number of
flowers and tall plant height (PC1); and, on species with longer styles
and higher levels of autonomous selfing (PC3). Overall, plant species
with small size, low nectar concentration and short styles showed lower
visitation rates than plant species with the opposed traits; species
with small amounts of floral rewards (pollen and nectar), low number of
flowers and small flowers had in general lower number of visiting
species (normalized degree); and species under 1.3 m, with high rewards
of pollen and nectar showed the highest levels of specialization. This
work highlights the need to consider essential plant reproductive and
floral traits to improve understanding of plant life strategies and
plant-pollinator interactions at large scales.

\vspace{3mm} \hrule
\vspace{5mm}

\emph{Keywords}: plant-pollinator networks \textbar{} floral traits
\textbar{} life history strategy \textbar{} functional traits

\bleft
\newpage

\section{INTRODUCTION}\label{introduction}

@@

\caption{\label{tab:unnamed-chunk-1}List of the 20 traits compiled in this study divided in quantitative and categorical traits. The different types of trait (vegetative ‘V’, floral ‘F’ and reproductive ‘R’), total records found for the 1,506 species and categories of the qualitative traits are also provided. }
\centering
\fontsize{10}{12}\selectfont
\begin{tabu} to \linewidth {>{}c>{\raggedright}X>{\centering}X>{}c>{\raggedright}X>{\raggedright}X>{\centering}X}
\toprule
\multicolumn{3}{c}{\textbf{Quantitative traits}} & \multicolumn{4}{c}{\textbf{Categorical traits}} \\
\cmidrule(l{3pt}r{3pt}){1-3} \cmidrule(l{3pt}r{3pt}){4-7}
\textbf{Type} & \textbf{Traits} & \textbf{Records} & \textbf{Type} & \textbf{Traits} & \textbf{Categories} & \textbf{Records}\\
\midrule
\textbf{V} & Plant height (m) & 1470 & \textbf{V} & Lifepan & \makecell[l]{Short-lived \\ Perennial} & 1466\\
\addlinespace
\textbf{F} & Flower width (mm) & 1472 & \textbf{V} & Life form & \makecell[l]{Herb \\ Shrub \\ Tree} & 1472\\
\addlinespace
\textbf{F} & Flower length (mm) & 1401 & \textbf{F} & Flower shape & \makecell[l]{Brush \\ Campanulate \\ Capitulum \\ Open \\ Papilionaceous \\ Tube} & 1458\\
\addlinespace
\textbf{F} & Inflorescence width (mm) & 1496 & \textbf{F} & Flower symmetry & \makecell[l]{Actinomorphic \\ Zygomorphic} & 1478\\
\addlinespace
\textbf{F} & Style length (mm) & 1497 & \textbf{F} & Nectar & \makecell[l]{Presence \\ Absence} & 1407\\
\addlinespace
\textbf{F} & Ovules per flower & 1457 & \textbf{R} & Autonomous selfing & \makecell[l]{None \\ Low \\ Medium \\ High} & 987\\
\addlinespace
\textbf{F} & Flowers per plant & 1468 & \textbf{R} & Compatibility system & \makecell[l]{Self-incomp. \\ Partially self-comp. \\ Self-comp.} & 1253\\
\addlinespace
\textbf{F} & Nectar ($\mu$l) & 485 & \textbf{R} & Breeding system & \makecell[l]{Hermaphrodite \\ Monoecious \\ Dioecious} & 1489\\
\addlinespace
\textbf{F} & Nectar (mg) & 415 & \textbf{} &  &  & \\
\addlinespace
\textbf{F} & Nectar concentration ($\%$) & 508 & \textbf{} &  &  & \\
\addlinespace
\textbf{F} & Pollen grains per flower & 298 & \textbf{} &  &  & \\
\addlinespace
\textbf{R} & Autonomous selfing (fruit set) & 992 & \textbf{} &  &  & \\
\bottomrule
\end{tabu}
\end{table}

@@

@@

\begin{figure}
\centering
\includegraphics{output/figures/unnamed-chunk-2-1.pdf}
\caption{\label{fig:unnamed-chunk-2}Life history strategies for 1,236 plant
species from 29 plant-pollinator networks studies across the first two
axes of variation from a phylogenetically informed principal component
analysis (pPCA). The solid arrows indicate the direction and weights of
the six quantitative traits (flower number, plant height, style length,
flower size, ovule number and autonomous selfing level) and the labelled
icons at their end represent the extreme form of the trait continuum.
The dashed lines indicate the opposed direction of trait variation and
the non-labelled icons at their end illustrate the other extreme of the
continuum.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{output/figures/unnamed-chunk-3-1.pdf}
\caption{\label{fig:unnamed-chunk-3}Location of the qualitative traits in
the trait space for traits that showed a statistical association with
the main axis of trait variation. These different traits included:
compatibility system (a), life form (b), lifespan (c), breeding system
(d), flower shape (e) and flower symmetry (f).}
\end{figure}

@@

\includegraphics{output/figures/unnamed-chunk-4-1.pdf}

>>>>

\documentclass[12pt,a4paper,]{article}
\usepackage[]{mathpazo}
\usepackage{setspace}
\setstretch{2}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{
            pdftitle={Trade-offs among plant floral and reproductive traits determine interactions with floral visitors},
            pdfauthor={Jose B. Lanuza1,2*; Romina Rader1; Jamie Stavert3; Liam K. Kendall4; Manu E. Saunders1; Ignasi Bartomeus2},
            colorlinks=true,
            linkcolor=RoyalBlue,
            citecolor=Blue,
            urlcolor=RoyalBlue,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

% Add Supplementary Tables and Figures
% Code from https://stackoverflow.com/a/51337664
\newcommand{\beginsupplement}{
  \setcounter{table}{0}  
  \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} 
  \renewcommand{\thefigure}{S\arabic{figure}}
}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{float}
\floatplacement{figure}{H}
\usepackage[skip=5pt]{caption}
\usepackage{float}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\title{Trade-offs among plant floral and reproductive traits determine
interactions with floral visitors}
\author{Jose B. Lanuza\textsuperscript{1,2}* \and Romina Rader\textsuperscript{1} \and Jamie Stavert\textsuperscript{3} \and Liam K. Kendall\textsuperscript{4} \and Manu E. Saunders\textsuperscript{1} \and Ignasi Bartomeus\textsuperscript{2}}
\date{}

@@

Plant life strategies are often delimited by vegetative and
physiological traits but little is known about how floral and
reproductive traits drive these strategies, and in turn shape plant
interactions with floral visitors. Here, we compiled 13 floral, 4
reproductive and 3 vegetative traits for 1,506 plant species from 28
plant-pollinator network studies across 18 different countries. We
investigated the associations among these traits, pollinator visitation
and the functional role of plant species within the networks
(interaction frequency, normalized degree and specialization). We found
that 51.8\% of trait variation was explained by two independent axes
that encompassed plant form and function. Specifically, the first axis
indicated the presence of a trade-off between flower number and flower
size (PC1, 26.72\%). The second axis indicated a trade-off for the level
of pollinator dependency (PC2, 25.08\%). Although the main axes of trait
variation did not fully explain pollinator visitation rates, different
plant life strategies were associated with visitation rates and
pollinator functional groups. Overall, the main traits that determined
plant species' functional roles were height, nectar concentration,
pollen grains per flower, number of ovules, style length, selfing level
and flower width. Our results highlight the need to consider plant
reproductive and floral traits to improve understanding of plant life
strategies and plant-pollinator interactions at broader spatial scales.

\vspace{5mm} \bleft

There is an astonishing diversity of floral structures and plant
reproductive strategies among flowering plants (Barrett, 2002; Schiestl
\& Johnson, 2013), which have long been of interest to pollination
biologists in terms of their relevance to plant-pollinator interactions.
However, most studies that have explored reproductive (e.g., mating and
compatibility systems) and floral trait (e.g., flower size or nectar
provision) variation have concentrated on the individual or community
level and thus, broader macroecological patterns remain poorly
investigated (Carvalheiro et al., 2014; Baude et al., 2016; Munoz et
al., 2016; Moeller et al., 2017; Grossenbacher et al., 2017). Indeed,
studies depicting species' life history strategies generally focus on
vegetative traits and rarely consider reproductive traits
(Salguero-Gómez et al., 2016; Rüger et al., 2018). As a consequence, a
unified framework that explores the compromises among floral traits and
their relevance to plant life strategies is currently lacking (Roddy et
al., 2021). At the same time, there is growing interest in the
determinants of plant-pollinator interactions via trait-based approaches
(Fenster et al., 2004) and trait-matching analyses (Bartomeus et al.,
2016). However, floral traits have been overlooked beyond highly
specialised plant-pollinator systems (Dellinger, 2020; Roddy et al.,
2021) and the role of plant reproductive biology remains little explored
in plant-pollinator interactions (but see Tur et al., 2013 and Devaux et
al., 2014).

With the recent availability of large trait databases, plant ecological
strategies are increasingly being examined (e.g., TRY, Kattge et al.,
2011 and COMPADRE, Salguero‐Gómez et al., 2015), and are facilitating
the identification of global patterns and constraints of plant form and
function (Salguero‐Gómez et al., 2015; Díaz et al., 2016; Carmona et
al., 2021). However, the main focus has been on vegetative traits such
as leaf (Wright et al., 2004) or wood (Chave et al., 2009) trade-offs
with little or no attention given to reproductive and floral traits
(E‐Vojtkó et al., 2020), also critical to plant form and function. For
instance, short lived versus perennial species tend to have low versus
high levels of outcrossing, respectively (Barrett, 2003; Moeller et al.,
2017). Further, outcrossing levels are positively correlated with flower
size (Goodwillie et al., 2010). In addition, the presence of costly
rewards (e.g., pollen or nectar) and showy flowers or floral displays
can only be understood through consideration of plant species' reliance
upon animal pollination (pollinator dependence) and their role in
attracting pollinators (Ollerton et al., 2011). Hence, exploring plant
life strategies with reproductive and floral trade-offs, in conjunction
with their pollinator dependence, is necessary for a balanced
understanding of plant economics.

Several studies have identified links between plant traits and
plant-pollinator network properties (Lázaro et al., 2008; Bartomeus,
2013; Carvalheiro et al., 2014). Moreover, plant traits can also define
species' network roles (e.g., specialists vs generalists). For example,
species that occupy trait space extremes are more likely to exhibit
higher specialization and be more reliant on trait-matching (Junker et
al., 2013; Coux et al., 2016). This morphological matching between plant
and floral visitors can determine plant-pollinator interactions, and
thus shape their interaction network structure (Stang et al., 2009;
Ibanez, 2012). Despite the increasing knowledge of the relevance of
traits on the species network roles, little is known about how plant
reproductive and floral traits determine plant species' network roles at
a macroecological scale.

Here, we explore the potential trade-offs among plant floral and
reproductive traits and how these influence the structure of
plant-pollinator networks. First, we identify the major axes of floral
and reproductive trait variation and trade-offs that determine plant
form and function. Second, we investigate how plant species' position in
trait-space influences interaction strength with different guilds of
floral visitors. Finally, we investigate how the main axes of trait
variation and individual traits influence plant species roles within
networks using complementary interaction network metrics (i.e.,
interaction strength, normalized degree and specialization).

\textbf{Plant strategies}

The phylogenetically informed principal component analysis (pPCA)
captured by the first two and three axes 51.8\% and 70.97\% of trait
variation, respectively (Figure 1 and Figure S5) and had a phylogenetic
correlation (\textbf{lambda}) of 0.76. The first principal component
(PC1) represented 26.72\% of the trait variation and indicated a
trade-off between flower number and flower size. We refer to this axis
as the `flower number - flower size trade-off as already described in
previous studies (Sargent et al., 2007; Kettle et al., 2011). Hence, one
end of the spectrum comprised species with high investment in flower
number and plant height but small flower size, short style length and
low ovule number. The other end of this spectrum comprised species that
were short in height and invested in large flowers, long styles, many
ovules, but few flowers. The main contributing traits to PC1 were plant
height, flower number, ovule number and flower size (loadings
\textgreater{} \textbar{}0.5\textbar{}; Table S3) but style length also
contributed moderately on PC1 (loading = -0.33). The second principal
component (PC2) represented 25.05\% of the trait variation and indicated
a trade-off between low and high pollinator dependence. We refer to this
axis as the `pollinator dependence trade-off'. The main driver of trait
variation on PC2 was autonomous selfing (loading = 0.85) but the other
traits (except ovule number) also made moderate contributions (loadings
from 0.27 to 0.4; Table S3). We found that high pollinator dependence
was associated with larger and higher numbers of flowers, greater plant
height and longer styles. In contrast, species with high levels of
autonomous selfing tended to have fewer and smaller flowers, had shorter
styles and were shorter in height. Further, PC3 explained a considerable
amount of trait variability (19.17\%) and the main contributors to this
axis were style length (loading = -0.66) and the degree of autonomous
selfing (loading = -0.51). The remaining traits, apart from ovule
number, were moderately correlated to changes on PC3 (loadings from
-0.23 to -0.46; Table S3). Thus, because style length was correlated
with all traits on PC3 and was the main driver of variation, we refer to
this axis as the `style length trade-off'. Further, the pPCA with the
subset of species that had nectar and pollen quantity data showed that
nectar quantity (microlitres of nectar per flower) was positively
associated with flower size, style length and ovule number (PC1,
23.40\%); and pollen quantity (pollen grains per flower) was positively
correlated with flower number and plant height and negatively associated
with autonomous selfing (PC2, 21.67\%; Figure S6). This pPCA explained
similar variance with the first two principal components (45.07\%) and
similar associations of traits despite some variability in the loadings
(Table S4).

\begin{figure}
\centering
\includegraphics{output/figures/unnamed-chunk-1-1.pdf}
\caption{\label{fig:unnamed-chunk-1}Figure 1. Life history strategies for
1,236 plant species from 28 plant-pollinator networks studies across the
first two axes of trait variation from a phylogenetically informed
principal component analysis (pPCA). The solid arrows indicate the
direction of the different quantitative traits (flower number, plant
height, style length, flower size, ovule number and level of autonomous
selfing) in the 2-d plane. The length of the arrows indicate the weight
of the variables on each principal component and the labelled icons at
their end represent the extreme form of the trait continuum. The dashed
lines show the opposed direction of trait variation and the non-labelled
icons at their end illustrate the opposing extreme of the continuum.}
\end{figure}

We found that most categorical traits were statistically associated with
the first two axes of trait variation (Figure 2 and Table S2). Flower
symmetry, which was only associated with PC2 (Sum of squares = 8.51,
F-value = 14.72, p \textless{} 0.01 ), and nectar provision, which was
independent of PC1 and PC2 (PC1: Sum of squares = 0.37, F-value = 0.29 ,
p = 0.59; PC2: Sum of squares = 0.83, F-value = 1.43, p = 0.23) showed
lack of statistical association. In addition, we found statistical
differences (Tukey test) between the different levels of categorical
traits in the trait space (Figure S7). Regarding self compatibility, we
found larger differences on PC2 (i.e., species with unisexual flowers
that were self incompatibile were statistically differentiated from
species with partial or full self compatibility; Figure S7 A and B).
Life forms differed statistically across both axes of trait variation
and followed a gradient of larger life forms (trees and shrubs) with
higher pollinator dependence to smaller ones (herbs) with lower
pollinator dependence (Figure S7 C and D). Consequently, lifespan also
followed this gradient but perennial and short lived species only
differed statistically on PC2 (Figure S7 E and F). Species with
unisexual flowers (monoecious and dioecious) were clustered on both
extremes of the first two principal components and had the highest
pollinator dependence and highest number of flowers (Figure S7 G and H).
Moreover, we found that the campanulate and capitulum flower shapes were
differentiated from tube, papilionaceous, open and brush shapes in the
trait space. The former morphologies had larger flowers and greater
pollinator dependence, while the latter had higher flower number and
greater autonomous selfing (Figure S7 I and J). Regarding flower
symmetry, zygomorphic flowers were associated with lower levels of
pollinator dependence, whereas actinomorphic flowers had higher levels
of pollinator dependence (Figure S7 K and L).

\blandscape

\begin{figure}[H]
\includegraphics{output/figures/unnamed-chunk-2-1} \caption{Location of the different qualitative traits that showed statistical association with the first two axes of trait variation. These different traits included: compatibility system (a), life form (b), lifespan (c), breeding system (d), flower shape (e) and flower symmetry (f).}\label{fig:unnamed-chunk-2}
\end{figure}

\elandscape

\textbf{Phylogenetic signal of traits}

\subsubsection{OLD}\label{old}

@@

\caption{\label{tab:unnamed-chunk-3}List of the 20 traits compiled in this study divided in quantitative and categorical traits. The different types of trait (vegetative ‘V’, floral ‘F’ and reproductive ‘R’), total records found for the 1,506 species and categories of the qualitative traits are also provided. }
\centering
\fontsize{10}{12}\selectfont
\begin{tabu} to \linewidth {>{}c>{\raggedright}X>{\centering}X>{}c>{\raggedright}X>{\raggedright}X>{\centering}X}
\toprule
\multicolumn{3}{c}{\textbf{Quantitative traits}} & \multicolumn{4}{c}{\textbf{Categorical traits}} \\
\cmidrule(l{3pt}r{3pt}){1-3} \cmidrule(l{3pt}r{3pt}){4-7}
\textbf{Type} & \textbf{Traits} & \textbf{Records} & \textbf{Type} & \textbf{Traits} & \textbf{Categories} & \textbf{Records}\\
\midrule
\textbf{V} & Plant height (m) & 1470 & \textbf{V} & Lifepan & \makecell[l]{Short-lived \\ Perennial} & 1466\\
\addlinespace
\textbf{F} & Flower width (mm) & 1472 & \textbf{V} & Life form & \makecell[l]{Herb \\ Shrub \\ Tree} & 1472\\
\addlinespace
\textbf{F} & Flower length (mm) & 1401 & \textbf{F} & Flower shape & \makecell[l]{Brush \\ Campanulate \\ Capitulum \\ Open \\ Papilionaceous \\ Tube} & 1458\\
\addlinespace
\textbf{F} & Inflorescence width (mm) & 1496 & \textbf{F} & Flower symmetry & \makecell[l]{Actinomorphic \\ Zygomorphic} & 1478\\
\addlinespace
\textbf{F} & Style length (mm) & 1497 & \textbf{F} & Nectar & \makecell[l]{Presence \\ Absence} & 1408\\
\addlinespace
\textbf{F} & Ovules per flower & 1457 & \textbf{R} & Autonomous selfing & \makecell[l]{None \\ Low \\ Medium \\ High} & 987\\
\addlinespace
\textbf{F} & Flowers per plant & 1468 & \textbf{R} & Compatibility system & \makecell[l]{Self-incomp. \\ Partially self-comp. \\ Self-comp.} & 1253\\
\addlinespace
\textbf{F} & Nectar ($\mu$l) & 531 & \textbf{R} & Breeding system & \makecell[l]{Hermaphrodite \\ Monoecious \\ Dioecious} & 1489\\
\addlinespace
\textbf{F} & Nectar (mg) & 455 & \textbf{} &  &  & \\
\addlinespace
\textbf{F} & Nectar concentration ($\%$) & 558 & \textbf{} &  &  & \\
\addlinespace
\textbf{F} & Pollen grains per flower & 533 & \textbf{} &  &  & \\
\addlinespace
\textbf{R} & Autonomous selfing (fruit set) & 992 & \textbf{} &  &  & \\
\bottomrule
\end{tabu}
\end{table}

@@

\section{RESULTS}\label{results-1}

@@

\begin{figure}
\centering
\includegraphics{output/figures/unnamed-chunk-4-1.pdf}
\caption{\label{fig:unnamed-chunk-4}Life history strategies for 1,236 plant
species from 29 plant-pollinator networks studies across the first two
axes of variation from a phylogenetically informed principal component
analysis (pPCA). The solid arrows indicate the direction and weights of
the six quantitative traits (flower number, plant height, style length,
flower size, ovule number and autonomous selfing level) and the labelled
icons at their end represent the extreme form of the trait continuum.
The dashed lines indicate the opposed direction of trait variation and
the non-labelled icons at their end illustrate the other extreme of the
continuum.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{output/figures/unnamed-chunk-5-1.pdf}
\caption{\label{fig:unnamed-chunk-5}Location of the qualitative traits in
the trait space for traits that showed a statistical association with
the main axis of trait variation. These different traits included:
compatibility system (a), life form (b), lifespan (c), breeding system
(d), flower shape (e) and flower symmetry (f).}
\end{figure}

@@

\includegraphics{output/figures/unnamed-chunk-6-1.pdf}
